Heuristics and Really Hard Instances for Subgraph Isomorphism Problems

### 2016-02-01 paper/paper.tex
Space, etc

The \emph{non-induced subgraph isomorphism problem} is to find an injective mapping from a given
pattern graph to a given target graph which preserves adjacency---in essence, we are ``finding a
copy of'' the pattern inside the target. The \emph{induced} variant of the problem additionally
requires that the mapping preserve non-adjacency, so there are no ``extra edges'' in the copy of the
pattern that we find. We illustrate both variants in \cref{figure:sip}.
Despite these problems being \NP-complete, modern practical subgraph isomorphism algorithms can
handle problem instances with many hundreds of vertices in the pattern graph, and up to ten thousand
vertices in the target graph \citep{Cordella:2004,Solnon:2010,Audemard:2014,McCreesh:2015}, leading
to successful application in areas such as computer
vision \citep{Damiand:2011,Solnon:2015}, biochemistry \citep{Giugno:2013,Carletti:2015}, and pattern
recognition \citep{Conte:2004}.

\begin{figure}[b]
    \centering
    \begin{tikzpicture}[scale=0.5]
        \begin{scope}
            \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (Na) at (1,  0) {1};
            \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (Nb) at (1, -1.5) {2};
            \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (Nc) at (0, -3) {3};
            \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (Nd) at (2, -3) {4};

            \draw [ultra thick] (Na) -- (Nb);
            \draw [ultra thick] (Nb) -- (Nc);
            \draw [ultra thick] (Nc) -- (Nd);
            \draw [ultra thick] (Nb) -- (Nd);

            \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (N1) at (5.5,  0) {1};
            \node [draw, circle, fill=white, inner sep=1.5pt] (N2) at (7.5,  0) {2};
            \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (N3) at (5.5, -1.5) {3};
            \node [draw, circle, fill=white, inner sep=1.5pt] (N4) at (7.5, -1.5) {4};
            \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (N5) at (5.5, -3) {5};
            \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (N6) at (7.5, -3) {6};

            \draw [thick, color=uofgsandstone!50] (N1) -- (N2);
            \draw [ultra thick] (N1) -- (N3);
            \draw [thick, color=uofgsandstone!50] (N1) -- (N4);
            \draw [thick, color=uofgsandstone!50] (N2) -- (N4);
            \draw [ultra thick] (N3) -- (N5);
            \draw [ultra thick] (N3) -- (N6);
            \draw [thick, color=uofgsandstone!50] (N4) -- (N6);
            \draw [ultra thick] (N5) -- (N6);
            \draw [thick, color=uofgsandstone!50] (N2) to [in=45, out=315] (N6);
            \draw [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);

            \node [draw, circle, fill=white, inner sep=1.5pt] (M1) at (-5.5,  0) {1};
            \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (M2) at (-3.5,  0) {2};
            \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (M3) at (-5.5, -1.5) {3};
            \node [draw, circle, fill=white, inner sep=1.5pt] (M4) at (-3.5, -1.5) {4};
            \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (M5) at (-5.5, -3) {5};
            \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (M6) at (-3.5, -3) {6};

            \draw [thick, color=uofgsandstone!50] (M1) -- (M2);
            \draw [thick, color=uofgsandstone!50] (M1) -- (M3);
            \draw [thick, color=uofgsandstone!50] (M1) -- (M4);
            \draw [thick, color=uofgsandstone!50] (M2) -- (M4);
            \draw [ultra thick] (M3) -- (M5);
            \draw [ultra thick] (M3) -- (M6);
            \draw [thick, color=uofgsandstone!50] (M4) -- (M6);
            \draw [ultra thick] (M5) -- (M6);
            \draw [ultra thick] (M2) to [in=45, out=315] (M6);
            \draw [thick, color=uofgsandstone!50] (M1) to [in=135, out=225] (M5);

            \node [anchor=center, font=\Large] (A1) at (-1.5, -1.5) { $\hookleftarrow$ };
            \node [anchor=center, font=\Large] (A2) at ( 3.5, -1.5) { $\rightarrowtail$ };
        \end{scope}
    \end{tikzpicture}

    \caption{There is a non-induced isomorphism from the pattern graph, in the centre, to the target
    on the right, mapping vertex 1 to 1, 2 to 3, 3 to 5 and 4 to 6. This is not an induced
    isomorphism, since there is an edge between 1 and 5 in the target but not between 1 and 3 in the
    pattern. The mapping from the pattern to the (same) target on the left, sending 1 to 2, 2 to 6,
    3 to 5 and 4 to 3, is both non-induced and induced.}
    \label{figure:sip}
\end{figure}

@@

Constrainedness, denoted $\kappa$, is an alternative measure of difficulty designed to generalise
hardness parameters across different combinatorial problems \citep{Gent:1996:Kappa}. A problem with
$\kappa < 1$ is said to be underconstrained, and is likely to be satisfiable; a problem with $\kappa
> 1$ is overconstrained, and is likely to be unsatisfiable. Empirically, problems with $\kappa$
close to 1 are hard, and problems where $\kappa$ is very small or very large are usually easy. By
handling injectivity as a restriction on the size of the state space rather than as a constraint, we
derive \[ \kappa = 1 - \frac{\log t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 -
d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}}}{\log t^{\underline{p}}}\] for induced isomorphisms, which
we plot on the sixth row of \cref{figure:induced}. We see that constrainedness predicts that the
central region will still be relatively difficult for larger pattern graphs: although the problem is
overconstrained, it is less overconstrained than in the regions the Glasgow and LAD solvers found easy.
Thus it seems that rather than just being a unification of existing generalised heuristic
techniques, constrainedness also gives a better predictor of difficulty than proximity to a phase
transition.  Unfortunately, constrainedness does not help us with heuristics: minimising
constrainedness gives the same predictions as maximising the expected number of solutions.

>>>>

The \emph{non-induced subgraph isomorphism problem} is to find an injective mapping from a given
pattern graph to a given target graph which preserves adjacency---in essence, we are ``finding a
copy of'' the pattern inside the target. The \emph{induced} variant of the problem additionally
requires that the mapping preserve non-adjacency, so there are no ``extra edges'' in the copy of the
pattern that we find. % We illustrate both variants in \cref{figure:sip}.
Despite these problems being \NP-complete, modern practical subgraph isomorphism algorithms can
handle problem instances with many hundreds of vertices in the pattern graph, and up to ten thousand
vertices in the target graph \citep{Cordella:2004,Solnon:2010,Audemard:2014,McCreesh:2015}, leading
to successful application in areas such as computer
vision \citep{Damiand:2011,Solnon:2015}, biochemistry \citep{Giugno:2013,Carletti:2015}, and pattern
recognition \citep{Conte:2004}.

%%% \begin{figure}[b]
%%%     \centering
%%%     \begin{tikzpicture}[scale=0.5]
%%%         \begin{scope}
%%%             \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (Na) at (1,  0) {1};
%%%             \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (Nb) at (1, -1.5) {2};
%%%             \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (Nc) at (0, -3) {3};
%%%             \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (Nd) at (2, -3) {4};
%%% 
%%%             \draw [ultra thick] (Na) -- (Nb);
%%%             \draw [ultra thick] (Nb) -- (Nc);
%%%             \draw [ultra thick] (Nc) -- (Nd);
%%%             \draw [ultra thick] (Nb) -- (Nd);
%%% 
%%%             \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (N1) at (5.5,  0) {1};
%%%             \node [draw, circle, fill=white, inner sep=1.5pt] (N2) at (7.5,  0) {2};
%%%             \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (N3) at (5.5, -1.5) {3};
%%%             \node [draw, circle, fill=white, inner sep=1.5pt] (N4) at (7.5, -1.5) {4};
%%%             \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (N5) at (5.5, -3) {5};
%%%             \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (N6) at (7.5, -3) {6};
%%% 
%%%             \draw [thick, color=uofgsandstone!50] (N1) -- (N2);
%%%             \draw [ultra thick] (N1) -- (N3);
%%%             \draw [thick, color=uofgsandstone!50] (N1) -- (N4);
%%%             \draw [thick, color=uofgsandstone!50] (N2) -- (N4);
%%%             \draw [ultra thick] (N3) -- (N5);
%%%             \draw [ultra thick] (N3) -- (N6);
%%%             \draw [thick, color=uofgsandstone!50] (N4) -- (N6);
%%%             \draw [ultra thick] (N5) -- (N6);
%%%             \draw [thick, color=uofgsandstone!50] (N2) to [in=45, out=315] (N6);
%%%             \draw [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
%%% 
%%%             \node [draw, circle, fill=white, inner sep=1.5pt] (M1) at (-5.5,  0) {1};
%%%             \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (M2) at (-3.5,  0) {2};
%%%             \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (M3) at (-5.5, -1.5) {3};
%%%             \node [draw, circle, fill=white, inner sep=1.5pt] (M4) at (-3.5, -1.5) {4};
%%%             \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (M5) at (-5.5, -3) {5};
%%%             \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (M6) at (-3.5, -3) {6};
%%% 
%%%             \draw [thick, color=uofgsandstone!50] (M1) -- (M2);
%%%             \draw [thick, color=uofgsandstone!50] (M1) -- (M3);
%%%             \draw [thick, color=uofgsandstone!50] (M1) -- (M4);
%%%             \draw [thick, color=uofgsandstone!50] (M2) -- (M4);
%%%             \draw [ultra thick] (M3) -- (M5);
%%%             \draw [ultra thick] (M3) -- (M6);
%%%             \draw [thick, color=uofgsandstone!50] (M4) -- (M6);
%%%             \draw [ultra thick] (M5) -- (M6);
%%%             \draw [ultra thick] (M2) to [in=45, out=315] (M6);
%%%             \draw [thick, color=uofgsandstone!50] (M1) to [in=135, out=225] (M5);
%%% 
%%%             \node [anchor=center, font=\Large] (A1) at (-1.5, -1.5) { $\hookleftarrow$ };
%%%             \node [anchor=center, font=\Large] (A2) at ( 3.5, -1.5) { $\rightarrowtail$ };
%%%         \end{scope}
%%%     \end{tikzpicture}
%%% 
%%%     \caption{There is a non-induced isomorphism from the pattern graph, in the centre, to the target
%%%     on the right, mapping vertex 1 to 1, 2 to 3, 3 to 5 and 4 to 6. This is not an induced
%%%     isomorphism, since there is an edge between 1 and 5 in the target but not between 1 and 3 in the
%%%     pattern. The mapping from the pattern to the (same) target on the left, sending 1 to 2, 2 to 6,
%%%     3 to 5 and 4 to 3, is both non-induced and induced.}
%%%     \label{figure:sip}
%%% \end{figure}

@@

Constrainedness, denoted $\kappa$, is an alternative measure of difficulty designed to generalise
hardness parameters across different combinatorial problems \citep{Gent:1996:Kappa}. A problem with
$\kappa < 1$ is said to be underconstrained, and is likely to be satisfiable; a problem with $\kappa
> 1$ is overconstrained, and is likely to be unsatisfiable. Empirically, problems with $\kappa$
close to 1 are hard, and problems where $\kappa$ is very small or very large are usually easy. By
handling injectivity as a restriction on the size of the state space rather than as a constraint, we
derive \[ \kappa = 1 - \frac{\log \left( t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 -
d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} \right)}{\log t^{\underline{p}}}\] for induced isomorphisms, which
we plot on the sixth row of \cref{figure:induced}. We see that constrainedness predicts that the
central region will still be relatively difficult for larger pattern graphs: although the problem is
overconstrained, it is less overconstrained than in the regions the Glasgow and LAD solvers found easy.
Thus it seems that rather than just being a unification of existing generalised heuristic
techniques, constrainedness also gives a better predictor of difficulty than proximity to a phase
transition.  Unfortunately, constrainedness does not help us with heuristics: minimising
constrainedness gives the same predictions as maximising the expected number of solutions.
